---
title: "Module 8 R demo"
author: "Profs Caetano & Schwartz"
date: "March 14, 2022"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load penguins data
```{r, include=FALSE}
library(tidyverse);  
library(palmerpenguins); # https://allisonhorst.github.io/palmerpenguins/articles/intro.html
# https://education.rstudio.com/blog/2020/07/palmerpenguins-cran/
```



```{r}
# Let's take a first look at the data
glimpse(penguins);
```


For this demo, we'll be trying to predict the length of a penguin's bill using a linear regression model.

```{r, fig.width=5, fig.height=2.5}
# First step let's look at the association between bill depth and bill length

# Let's get rid of the warning

# Now we can fit a linear regression model to start exploring this association more deeply

```

Do you think this fitted linear model is effectively representing the association between bill depth and bill length? What other variable could we add to this model? 
```{r}
#glimpse(penguins_clean)

# New variable to add to the model: 

library(broom)

# What about an interaction term?


## go to pollev.com/sta to make a prediction using this model!
```

# Comparing the prediction accuracy of multiple models

```{r}
# Set up


# Create training dataset


# Testing dataset includes all observations NOT in the training data



# Fit models to training data



# Make predictions for testing data using training model


# Make predictions for training data using training model


# Calculate RMSE for testing data

# Calculate RMSE for training data

```


### What does it mean if the RMSE based on test data is *smaller* than the RMSE based on the training data?






